Title goes here

### 2012-06-26 paper.tex
Wrote motivation, commit before david's update

\begin{abstract}
%\boldmath
The abstract goes here.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

\section{Motivation}

@@

% use section* for acknowledgement
\section*{Acknowledgment}
This work is funded by the IRCSET Enterprise Partnership Scheme along with IBM Research Ireland.

@@

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

>>>>

\begin{abstract}
%\boldmath
the abstract goes here.

\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

\section{Motivation}
In the continued race for improvement of multi core and many core architectures, computer networks are becoming a fundamental area for performance bottleneck especially in systems which are used for distributed processing of parallel applications. In order to handle the requirements posed by these applications the interconnects are becoming more and more complex. Such systems utilize a wide range of custom interconnects ranging from infiniband, myrinet to the more readily available ones such as 1Gbe and 10Gbe Ethernet standards and sophisticated routers and switches to handle the added challenge in routing data. While a small group of the specialized systems employ custom routers and topology configurations the majority of them especially those that use 1Gbe or 10Gbe use the commercial routers. In addition the growing complexity of routing protocols has resulted in advanced switches such as OpenVSwitch \cite{openvswitch}, etc., has resulted in these routers being upgraded frequently to keep up with the growing requirements. Software routers such as click modular router have been gaining popularity due to their flexible nature and cheaper alternative to the more expensive commercial routers and are widely used to prototype newer standards. In using software routers in the above scenario it would be possible to not only benefit from the newer developments in switching and routing protocols but it would require little effort to implement it in existing systems. One might argue that it would be a challenge to handle such complex workloads in software routers but with the presence of multicore and many core architectures and stream processors it is feasible to parallelize the packet processing workloads to improve the overall performance of the system. Several research projects are working on improving the performance by parallelizing the packet processing tasks on these architectures \cite{routebricks} \cite{packetshader}.

In a computing cluster which is used in the execution of parallel applications it is common to dedicate a core or processor in each node for operating system related tasks in order to minimize its effects on the application. Similarly, it should also be possible to utilize an additional core or processor to perform packet processing operations. While this is would be a challenging scenario it does provide the added benefits of a cheaper alternative to commercial routers. Also in bringing the network system into a software controlled domain it should be possible to perform additional optimizations from the application such as prioritized data flow, intelligent routing, etc., that would otherwise be difficult to implement due to the abstraction of the network system.

The focus of this paper is the development of a parallel processing engine that could be integrated with the application and can perform both the actual computations and the packet processing related tasks in parallel. The entire system can be divided into key parts as follows,

\begin{itemize}

\item A suitable methodology by which we can access the data in the userspace without interacting with the operating system. To do this (1)We need to read the data from the Network Interface Controller (NIC) before the OS and (2)Using this, establish a efficient manner for sending and receiving the data.(Section \ref{sec:userspace-packet})

\item Given a graph that represents the workload to be done on the packet flow, we divide it into tasks that can be executed in parallel and determine the regions that could be used to extract data and task parallelism.(Section \ref{sec:workload})

\item The parallel processing engine that utilizes the above and does the actual computations on the multi core or many core processor. It also handles load balancing and scheduling of the work across the available computing resources. (Section \ref{sec:ppe})

\end{itemize}

We concentrate more on the framework that can integrate the packet processing and application tasks into a parallel processing engine and the extraction of parallelism from the packet processing related tasks to improve the overall performance. Given an application graph and its partition consisting of parallel tasks we the combine it with the packet processing operations. In addition, Section \ref{sec:} we show the challeneges involved in using a system consisting of threads and process that are used to actually perform the work. The extent to which we extract parallelism ranges not only to the packet tasks but to the low level access of data I/O to the network interface hardware through certain additions to the netmap APIs. Finally in Section \ref{sec: } we present certain pseudo applications that involves operations such as compression and encryption to evaluate our parallel processing engine.

An important part of our system is the tasks related to packets such as routing, forwarding, etc., Since our design focuses more on the aspects of the parallel framework we utilize the click modular router for the packet processing tasks. Click modular router \cite{click} is a commonly used software router that includes a library of elements which is used in building flexible and reconfigurable routers. It offers a platform that is flexible for rapid development of newer network protocols and achieves good performance. An improved version known as SMP Click \cite{smpclick} also supports thread based parallel processing of network related tasks. Until recently click's performance in the user space was not as good as its kernel space equivalent. This is mainly due to its extensive use of polling driver in the kernel space. Netmap I/O \cite{netmap} provides a set of APIs that could directly access the packets from the NIC and it helped in the recent improvement of click's user space performance.

Click dynamically constructs routers in which the configuration file is passed as a parameter for the kernel module or user space binary. As a result most of the actual components are dynamically allocated and connected at run time. While this may be an advantage considering the alternate is to recompile the linux kernel, it does act as significant bottleneck in the case of user space. Moreover some of the restrictions on how the configuration can be parallelized limits the actual performance that could be achieved in combination with netmap. By integrating the ability to access click's elements within our parallel processing engine we were able to scale significantly better. This does require a limited amount of modifications to some of click's libraries and elements. In generating the actual code statically, compilers such as gnu c, etc., can perform optimizations such as loop fusions, in-lining, dead code optimization etc., which were not possible previously.

Our framework primarily focuses on user space and netmap APIs forms a crucial part in this.

@@

% use section* for acknowledgement
\section*{Acknowledgment}
This work is funded by the IRCSET Enterprise Partnership Scheme along with IBM Research, Ireland.

@@
